---
title: "Vizualization Part 2"
author: "Diana Ballesteros"
date: "10/2/2018"
output: github_document
---

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

devtools::install_github("thomasp85/patchwork")

```

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```

Data import. 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Start a plot 

First Scatterplot! 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

You can add titles to your graphs, very important esp for others to see 
In ggplot, there are multiple ways to do things. (xlab() or ylab)
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) 
```

Tick marks and labels...
scale_x_continuous allows you to do many things to the aesthetics of the plot. 
Breaks allows you to add tick marks in certain places 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  scale_x_continuous(breaks = c(-15, 0, 15), 
                     labels = c("-15ยบ C", "0", "15"))
```

trans = allows you to transform your values 
```{r}

## limits allows you to put a limit on your plot. Think of like a range. 
## You can also add the scale_y_continuos to the y 
## position = "right" allows you to move the tick marks to the right. 
## trans = "sqrt" allows you to transformt the data 
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  scale_x_continuous(breaks = c(-15, 0, 15), 
                     labels = c("-15ยบC", "0", "15"),
                     limits = c(-20, 30)) + 
  scale_y_continuous(trans = "sqrt", 
                     position = "right")
```

## Colors and Themes
Adjsut color...
scale_color_discrete is the same as scale_color_hue
name = "location" allows you to change the name of the title of the legend key 
h = is the hue 
l = luminosit 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  scale_color_hue(name = "Location", 
                  h = c(100, 350), 
                  l = 75)
```

Viridis
by default, viridus will do a continous variable, so need to tell it is discrete
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE) 
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme(legend.position = "bottom")
```

## Theme 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

```{r}
## Add the legend to the bottom, where you add the theme - "bottom" matters. 
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  ggthemes::theme_excel() + 
  theme(legend.position = "bottom")
```

## Learning Assessment 
Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot.
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) + 
  geom_smooth(se = FALSE) + 
  labs(
    title = "Temperature vs. Date plot",
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

## Arguments to 'geom_*'
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

```{r}
## You can also do the same thing by inserting facet_grid 
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

## Patchwork 
```{r}
library(patchwork)
tmax_tmin_p = ggplot(weather_df, aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

## Doing this will combine all of yourplots that are fundamentally different
(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  facet_grid(~name) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

## Data Manipulation 

```{r}
ggplot(weather_df, aes(x = name, y = tmax, fill = name)) + 
  geom_violin()
```

```{r}
## It now puts it a new order of my chosing; allows you to rearrange the graphs in the order you like 
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```


```{r}

## forcats::fct_reorder allows you rearrange according to a new factor of your choosing
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

Advanced tidying...
```{r}

## Data manipulation vs. Data tidying at the same time 
weather_df %>%
  select(name, tmax, tmin) %>% 
  gather(key = observation, value = temp, tmax:tmin) %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```

## can't load becayse I don't have the data set in the data folder...
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  gather(key = visit, value = bdi, bdi_score_bl:bdi_score_12m) %>%
  separate(visit, into = c("remove_1", "remove_2", "visit"), sep = "_") %>%
  select(id, visit, everything(), -starts_with("remove")) %>%
  mutate(visit = replace(visit, visit == "bl", "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
  